The Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams

### 2016-09-21 apjRevised/response.txt

REVIEWER:

1) The authors calculated the linear growth rate by adopting the inappropriate relativistically drifting Maxwellian distribition function (Eqs. (1) and (3)) for the pair
production spectrum from the double photon annihilation process. But this Maxwellian distribution function does not result from the double photon annihilation process. The correct pair production spectrum has been calculated before in Eq. (154) of Schlickeiser et al. (2012, ApJ 758, 101) -- see also the approximations (3) and (5) in Schlickeiser, Krakau and Supsar (2013, ApJ 777, 49) -- and by Aharonian, Atoyan and Nagapetian (1983, Astrofizika (Astrophysics), 19, 1), depending on the incoming TeV gamma-ray enery spectrum and the EBL target photon spectrum. This results in a power law energy spectrum with a strong exponential cut-off at small electron energies, which is substantially different from a drifting Maxwellian. It is necessary for the authors to redo their linear instability analysis with the correct pair production energy spectrum.

RESPONSE:

We agree that a drifting Maxwellian is not the correct distribution function to model a pair beam. But we adopt it because:
1. It is simpler and allows us to calculate the kinetic instability exactly in the electrostatic approximation.
2. It has all the features that allows us the elucidate the physics (continuous spread of momenta in the parallel and perpendicular directions)
3. It is physically motivated — by this we mean that it is a distribution that can be realized in physical situation.

We added note that we use the Maxwell-Juttner distribution in the abstract. We have added an additional paragraph to the end of section 2.1 elucidating this choice and saying that a correct calculation involves using a distribution function calculated from the \gamma\gamma process such as the ones mentioned by the referee.  We also note that the analytic techniques that we used for calculating the drifting Maxwellian may be useful in the full calculation.

### 2016-09-21 apjRevised/ms.tex
updated and massaged the text

\newcommand{\mathsc}[1]{\text{\textsc{#1}}}

@@

\begin{abstract}
The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that they drive.  Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes.  We reproduce the well-known reactive and kinetic growth rates for both the beam plasma and oblique mode. We demonstrate for the oblique instability that there is broad spectrum of unstable modes that grow at the maximum rate for a wide range of beam temperatures and wavevector orientations relative to the beam.  We also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower Lorentz factor than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.  Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the both the kinetic oblique mode and the reactive beam-plasma mode.  These results are important in understanding how powerful plasma instabilities may sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space.
\end{abstract}

@@

{\bf While we have modeled the pair distribution function as a Maxwell-J{\"u}ttner distribution, the distribution function that are produced by VHEGR photo annihilation is different \citep[see for instance][]{2012ApJ...758..101S}.  In particular, the the parallel and perpendicular momentum spread will likely be different due to the distribution of VHEGR photons and their respective mean free paths.  However, a Maxwell-J{\"u}ttner distribution is still useful as it has been calculated without approximation previously by \citet{Bret-Grem-Beni:10}, allowing a point of comparison for our calculation using the electrostatic approximation (as mentioned below).  In addition, it possess a continuous (small) distribution of parallel and perpendicular momenta that allow us to elucidate the physics.  It also is sufficiently simple to allows us to calculate the kinetic instability exactly in the electrostatic approximation.  Finally, we note that the analytic methodology use to calculate the Maxwell-J{\"u}ttner distribution may be useful for the full calculation using the physical distribution function.}

>>>>

\usepackage{ulem}
\newcommand{\mathsc}[1]{\text{\textsc{#1}}}

@@

\begin{abstract}
The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
Here we  study {\bf \sout{the initial conditions of these e± beams and}} the linear phase of the plasma instabilities that these pair beams drive.  Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes{\bf,  assuming a Maxwell-J{\"u}ttner distribution for the pair beam}.  We reproduce the well-known reactive and kinetic growth rates for both the beam plasma and oblique mode. We demonstrate for the oblique instability that there is broad spectrum of unstable modes that grow at the maximum rate for a wide range of beam temperatures and wavevector orientations relative to the beam.  We also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower Lorentz factor than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.  Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the both the kinetic oblique mode and the reactive beam-plasma mode.  These results are important in understanding how powerful plasma instabilities may sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space.
\end{abstract}

@@

{\bf While we have modeled the pair distribution function as a Maxwell-J{\"u}ttner distribution, the physical distribution function that is produced by VHEGR photo annihilation is likely different \citep[see for instance][]{2012ApJ...758..101S}.  In particular, the parallel and perpendicular momentum spread will be influenced by the distribution of VHEGR photons and their respective mean free paths.  However, a Maxwell-J{\"u}ttner distribution is still useful. First, it also is sufficiently simple to allows us to calculate the kinetic instability exactly in the electrostatic approximation.  Second, its instability growth rates has been calculated without approximation previously by \citet{Bret-Grem-Beni:10}, allowing a point of comparison for our calculation using the electrostatic approximation (as mentioned below).  Third, it possess a continuous (small) distribution of parallel and perpendicular momenta that allow us to elucidate the physics.   Finally, the analytic methodology used to calculate the Maxwell-J{\"u}ttner distribution may be useful for the full calculation using the physical distribution function.}
